---
title: 'DATA 555: Flexdashboard'
author: "Afia Tyus"
date: "2025-03-07"
output:
  flexdashboard::flex_dashboard:
    storyboard: TRUE
social: menu
source_code: embed
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(pacman)
p_load(flexdashboard, dplyr, ggplot2, esquisse, plotly, DT, scales, readr, stringr)

here::i_am("Code/Flexdashboard_v2.Rmd")
data_filtered <- read_csv(here::here("Output/data_filtered_AT.csv"))
```

------------------------------------------------------------------------

### Boxplot 1 : Feeding Practices Known by Household Type

```{r}
# Create a new variable to seperate jitter points from boxplot
data_filtered$HH_type_jitter <- as.numeric(factor(data_filtered$Gendered_HH_Type)) + 0.25  # Nudges jitter to the right

data_filtered$program <- factor(data_filtered$program,
                               levels = c("TAKUNDA","TITUKULANE"),
                               labels = c("Takunda (Zimbabwe)", 
                                          "Titukulane (Malawi)")
                               )

boxplot_1A <- ggplot(data_filtered %>% filter(KFP_sum != 0), aes(x= Gendered_HH_Type, y = KFP_sum, fill = program)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5, width = 0.4, position = position_nudge(x = -0.25)) +  # Nudges boxplots left
    geom_jitter(aes(x = HH_type_jitter, text = paste("Program:", program, "<br>Feeding Practices Known: ", KFP_sum, "<br>Age: ", Age, "<br>Sex: ", Sex, "<br>Interventions participated in: ", interventions)), 
              width = 0.15, size = 2, alpha = 0.7)+
  scale_x_discrete(labels = label_wrap(20)) +  
  scale_fill_discrete(name = "Program")+
  labs(x = "Gendered Household Type",
       y = "Total Feeding Practices Known (count)",
       title = "Gendered Household Type and Feeding Practices Known",
       subtitle = "This graph excludes those knowing zero (0) feeding practices to provide a clearer picture of knowledge distribution") +
  theme_minimal()

note_1 <- "Note: This graph excludes those knowing zero (0) feeding practices to provide a clearer picture of knowledge distribution"

plotly_1A <- ggplotly(boxplot_1A, tooltip = "text") %>% 
    layout(margin = list(l = 50, r = 50, b = 100, t = 50),
         annotations = list(x = 1, y = -0.3, text = str_wrap(note_1, 100),
                            xref='paper', yref='paper', showarrow = F, 
                            xanchor='auto', yanchor='auto', xshift=100, yshift=-20,
                            font = list(size = 10)))

plotly_1A

```

------------------------------------------------------------------------

Key Takeaways:

This graph provides a view of the distribution of feeding practices known, separated by gendered household type. When viewing this widget (with zero's excluded), it becomes clear that the two household types do not appear to have significantly different numbers of feeding practices known. It does lead to the belief that those in the Titukulane program on average know fewer feeding practices, while those in Takunda know significantly more.

### Boxplot 2: Feeding Practices known and household type (including zeros)

```{r}
boxplot_1B <- ggplot(data_filtered, aes(x= Gendered_HH_Type, y = KFP_sum, fill = program)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5, width = 0.4, position = position_nudge(x = -0.25)) +  # Nudges boxplots left
    geom_jitter(aes(x = HH_type_jitter, text = paste("Program:", program, "<br>Feeding Practices Known: ", KFP_sum, "<br>Age: ", Age, "<br>Sex: ", Sex, "<br>Interventions participated in: ", interventions)), 
              width = 0.15, size = 2, alpha = 0.7)+
  scale_x_discrete(labels = label_wrap(20)) +  
  scale_fill_discrete(name = "Program")+
  labs(x = "Gendered Household Type",
       y = "Total Feeding Practices Known (count)",
       title = "Gendered Household Type and Feeding Practices Known") +
  theme_minimal()

note_2 <- "Note: This graph includes those knowing zero (0) feeding practices"

plotly_1B <- ggplotly(boxplot_1B, tooltip = "text") %>% 
    layout(margin = list(l = 50, r = 50, b = 100, t = 50),
         annotations = list(x = 1, y = -0.3, text = str_wrap(note_2, 100),
                            xref='paper', yref='paper', showarrow = F, 
                            xanchor='auto', yanchor='auto', xshift=100, yshift=-20,
                            font = list(size = 10)))

plotly_1B
```

------------------------------------------------------------------------

Key Takeaways:

This graph provides a view of the distribution of feeding practices known, separated by gendered household type. When viewing this widget (with zero's INCLUDED), we see that across both household types, participants are significantly more likely to know zero (0) feeding practices than one or more. Overall, both household types have similar distributions of feeding practices known, and the coloring of the dots shows that program may be a stronger driver of knowledge than gendered household type.

### Interactive Table

```{r}

# Rename columns with title case and add units in parentheses
data_table <- data_filtered %>%
  select(
    Program = program,
    District = District, 
    Cluster = Cluster, 
    Village = Village, 
    Ward = Ward,
    `Gendered Household Type` = Gendered_HH_Type,
    `Age (Years)` = Age,
    `Interventions Participated In` = interventions,
    `Sex` = Sex,
    `Feeding Practices Known` = KFP_sum,
    `Feeding Practices (binary)` = KFP,
    `Men (15 - 29 years)` = Men_15_29,
    `Men (30+ years)` = Men_30,
    `Women (15 - 29 years)` = Female_15_29,
    `Women (30+ years)` = Female_30
  )

# Create interactive DataTable
table_1 <- datatable(data_table, 
          options = list(
            pageLength = 10,       # Show 10 rows per page by default
            autoWidth = TRUE,      # Auto adjust column width
            searching = TRUE,      # Enable search box
            fillContainer = TRUE,
            scrollX = TRUE, # enabling horizontal scrolling
            scrollY = "400px", # enabling vertical scrolling
            dom = 'Blfrtip',       # Show filter dropdowns
            columnDefs = list(     
              list(orderable = TRUE, targets = '_all')  # Make all columns sortable
            )
          ),
          rownames = FALSE,        # Hide row numbers
          filter = "top"           # Add column filters
)

table_1

```

------------------------------------------------------------------------

Key Takeaways:

This interactive data table allows the program coordinators to view and sort through the cleaned and mapped dataset. These programs have never been analyzed together (to date, they’ve produced country specific reports), as a result being able to easily view the new variables will aid in the applications of the research.

### Barchart

```{r}

# Create a new variable to seperate jitter points from boxplot
data_filtered$HH_type_jitter <- as.numeric(factor(data_filtered$Gendered_HH_Type)) + 0.25  # Nudges jitter to the right


# Calculate counts per program and KFP
data_with_count <- data_filtered %>%
  count(program, KFP) %>%
  mutate(text = paste("Program:", program, "<br>Feeding Practices Known:", KFP, "<br>Count:", n))

# Create the bar chart with the calculated counts
bar_chart <- ggplot(data = data_with_count, aes(fill = KFP, x = program, y = n)) +
  geom_bar(stat = "identity", position = "dodge", aes(text = text)) + 
  scale_x_discrete(labels = label_wrap(20)) +
  labs(x = "Program",
       y = "Participants (count)",
       title = "Program and Feeding Practices Known")

# Convert to plotly and specify the tooltip
ggplotly(bar_chart, tooltip = "text")

```

------------------------------------------------------------------------

Key Takeaways:

This barchart shows feeding practices known as a binary variable with participants either knowing 0 practices or 1 or more practices. Without the noise of gendered household type, this graph clearly shows that Takunda participants are more likely to know 1 or more feeding practices.

### Scatterplot

```{r}

interventions_v_KFP <- data_filtered %>% group_by(interventions, KFP_sum) %>%
  mutate(count = n()) %>%
  ggplot(aes(x = interventions, y = KFP_sum, color = program, size = count)) +
  geom_point(aes(text = paste("Program: ", program, "<br>Feeding Practices Known: ", KFP_sum, "<br>Interventions participated in: ", interventions, "<br>Participants: ", count))) +
  geom_smooth(method = "lm", se=FALSE) +
  labs(x = "Interventions Participated In",
       y = "Feeding Practices Known",
       title = "Intervention Participation and Feeding Practice Knowledge")

ggplotly(interventions_v_KFP, tooltip = "text")

```

------------------------------------------------------------------------

Key Takeaways:

This graph shows that interventions participated in has a slightly positive relationship with feeding practices known. Titukulane participants have a wider range of total interventions participated in (0 to 18), while Takunda participants participated in 0 to 10 interventions.
