---
title: "HTML Widgets"
author: "Afia Tyus"
date: "2025-03-05"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Setup

```{r}
# Loading packages and importing data
library(pacman)
p_load(dplyr, ggplot2, esquisse, plotly, DT, scales, readr, stringr)

here::i_am("07_Afia/HTML_Widget_v1.Rmd")
data_filtered <- read_csv(here::here("07_Afia/data_filtered_AT.csv"))

```

# Widget 1A: Boxplot of Feeding Practices Known

Using plotly, I’ll create a side by side boxplot comparing the knowledge
of feeding practices in each gendered household type

```{r}
# Create a new variable to seperate jitter points from boxplot
data_filtered$HH_type_jitter <- as.numeric(factor(data_filtered$Gendered_HH_Type)) + 0.25  # Nudges jitter to the right

data_filtered$program <- factor(data_filtered$program,
                               levels = c("TAKUNDA","TITUKULANE"),
                               labels = c("Takunda (Zimbabwe)", 
                                          "Titukulane (Malawi)")
                               )

boxplot_1A <- ggplot(data_filtered %>% filter(KFP_sum != 0), aes(x= Gendered_HH_Type, y = KFP_sum, fill = program)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5, width = 0.4, position = position_nudge(x = -0.25)) +  # Nudges boxplots left
    geom_jitter(aes(x = HH_type_jitter, text = paste("Program:", program, "<br>Feeding Practices Known: ", KFP_sum)), 
              width = 0.15, size = 2, alpha = 0.7)+
  scale_x_discrete(labels = label_wrap(20)) +  
  scale_fill_discrete(name = "Program")+
  labs(x = "Gendered Household Type",
       y = "Total Feeding Practices Known",
       title = "Gendered Household Type and Feeding Practices Known",
       subtitle = "This graph excludes those knowing zero (0) feeding practices to provide a clearer picture of knowledge distribution") +
  theme_minimal()

note_1 <- "Note: This graph excludes those knowing zero (0) feeding practices to provide a clearer picture of knowledge distribution"

plotly_1A <- ggplotly(boxplot_1A, tooltip = "text") %>% 
    layout(margin = list(l = 50, r = 50, b = 100, t = 50),
         annotations = list(x = 1, y = -0.3, text = str_wrap(note_1, 100),
                            xref='paper', yref='paper', showarrow = F, 
                            xanchor='auto', yanchor='auto', xshift=100, yshift=-20,
                            font = list(size = 10)))


plotly_1A
```

# Widget 1B: Boxplot of Feeding Practices Known

```{r}
boxplot_1B <- ggplot(data_filtered, aes(x= Gendered_HH_Type, y = KFP_sum, fill = program)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5, width = 0.4, position = position_nudge(x = -0.25)) +  # Nudges boxplots left
    geom_jitter(aes(x = HH_type_jitter, text = paste("Program:", program, "<br>Feeding Practices Known: ", KFP_sum)), 
              width = 0.15, size = 2, alpha = 0.7)+
  scale_x_discrete(labels = label_wrap(20)) +  
  scale_fill_discrete(name = "Program")+
  labs(x = "Gendered Household Type",
       y = "Total Feeding Practices Known",
       title = "Gendered Household Type and Feeding Practices Known") +
  theme_minimal()

note_2 <- "Note: This graph includes those knowing zero (0) feeding practices"

plotly_1B <- ggplotly(boxplot_1B, tooltip = "text") %>% 
    layout(margin = list(l = 50, r = 50, b = 100, t = 50),
         annotations = list(x = 1, y = -0.3, text = str_wrap(note_2, 100),
                            xref='paper', yref='paper', showarrow = F, 
                            xanchor='auto', yanchor='auto', xshift=100, yshift=-20,
                            font = list(size = 10)))

plotly_1B
```

# Widget 2: Bar graph

The second html widget will be a map with the median feeding practices
known coloring each location. Interactive Features :  Hover over each
location for the summary statistics of each location (mean, median, 1st
quartile, 3rd quartile, participants)  Zoom into each area to drill
down from country to ward, district, village, etc. Information that
users can extract from the visualization –  Trends in total number of
feeding practices known  Indication of potential regional differences
in knowledge  Distribution of feeding practice awareness

```{r}

# Create a new variable to seperate jitter points from boxplot
data_filtered$HH_type_jitter <- as.numeric(factor(data_filtered$Gendered_HH_Type)) + 0.25  # Nudges jitter to the right


bar_chart <- ggplot(data=data_filtered, aes(fill = KFP, x = program))+
  geom_bar(position="fill")+
  facet_wrap(~Gendered_HH_Type)+
  scale_x_discrete(labels = label_wrap(20))


ggplotly(bar_chart, tooltip = "text")
```

# Widget 3: Data Table

Using DT, I can also create an interactive data table that allows the
program coordinators to view and sort through the cleaned and mapped
dataset. These programs have never been analyzed together (to date,
they’ve produced country specific reports), as a result being able to
easily view the new variables (with a corresponding methodology) will
aid in the applications of the research. This would exclude information
like names and addresses.

```{r}

# Rename columns with title case and add units in parentheses
data_table <- data_filtered %>%
  select(
    Program = program,
    District = District, 
    Cluster = Cluster, 
    Village = Village, 
    Ward = Ward,
    `Gendered Household Type` = Gendered_HH_Type,
    `Age (Years)` = Age,
    `Interventions Participated In` = interventions,
    `Sex` = Sex,
    `Total Number of Feeding Practices Known` = KFP_sum,
    `Knows 1 or more feeding practices` = KFP,
    `Men in the household (between 15 and 29)` = Men_15_29,
    `Men in the household (over 30)` = Men_30,
    `Women in the household (between 15 and 29)` = Female_15_29,
    `Women in the household (over 30)` = Female_30
  )

# Create interactive DataTable
table_1 <- datatable(data_table, 
          options = list(
            pageLength = 10,       # Show 10 rows per page by default
            autoWidth = TRUE,      # Auto adjust column width
            searching = TRUE,      # Enable search box
            fillContainer = TRUE,
            scrollX = TRUE, # enabling horizontal scrolling
            scrollY = "400px", # enabling vertical scrolling
            dom = 'Blfrtip',       # Show filter dropdowns
            columnDefs = list(     
              list(orderable = TRUE, targets = '_all')  # Make all columns sortable
            )
          ),
          rownames = FALSE,        # Hide row numbers
          filter = "top"           # Add column filters
)

table_1


```

# Widget 4: Interactive Map

```{r}



```
